---
title: "Descriptive"
author: "Mary Ann Binuya"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
  pdf_document: default
---

# Aim
1. To create Table 1's and check distribution of outcomes/covariates.

# Set-up
```{r setup, include=FALSE}
library(dplyr)
library(table1)
library(survival)
library(survminer)
library(ggplot2)
```

# Load required datasets
```{r load}
load("WS_03_BOADICEA_output.RData")

   dat_all <- dat_full %>%
    mutate(predrisk_mod1 = BOA_5yrsrisk_Model1/100,
            predrisk_mod2 = BOA_5yrsrisk_Model2/100,
            predrisk_mod3a = BOA_5yrsrisk_Model3a/100,
            predrisk_mod3b = BOA_5yrsrisk_Model3b/100,
            predrisk_mod4 = BOA_5yrsrisk_Model4/100,
            predrisk_mod5a = BOA_5yrsrisk_Model5a/100,
            predrisk_mod5b = BOA_5yrsrisk_Model5b/100)
  
  rm(list=setdiff(ls(), c("dat_all")))
  
  nrow(dat_all) #n=1135
  
# Create subgroups
  dat_nodcis <- dat_all %>% filter(!(bc == 1 & omgedrag == "In situ")) #remove all DCIS from cases, n=1089
  
  dat_dcishg <- dat_all %>% filter(!(bc == 1 & omgedrag == "In situ" & diffgrad != 3)) #include high-grade DCIS as cases, n=1109
    
  dat_dcisall <- dat_all #include all DCIS as cases, n=1135
  
  rm(list=setdiff(ls(), c("dat_nodcis", "dat_dcishg", "dat_dcisall")))
```

# 1. Data: Only IBC included as cases; DCIS excluded
## Outcome and follow-up
```{r 1_outcome}
dat <- dat_nodcis

# Check outcome distributions
  table(dat$bc, useNA="always")
  
# Type of detection
  table(dat$scrrel[dat$bc==1], useNA="always")
  round(prop.table(table(dat$scrrel[dat$bc==1], useNA="always"))*100, 1)

# Follow-up time
  summary(dat$futime)

# Follow-up time by bc
  by(dat$futime, dat$bc, summary)


# Survival over time
  obs <- survfit(Surv(futime, bc)~ 1, weights = wt_KM, data=dat) 
  
  ## Risk tables
  tables <- ggsurvtable(obs, data = dat, break.time.by = 1)
  temp <- tables$risk.table$data
  risk_table <- round(data.frame(
    year = temp$time,
    n_risk = temp$n.risk,
    n_event = temp$n.event,
    cum_n_event = temp$cum.n.event,
    n_censor = temp$n.censor,
    cum_n_censor = temp$cum.n.censor),0)
  risk_table
  
    rm(temp, tables)

  ## Cumulative incidence (1-KM) plot
  ggsurvplot(
    obs, 
    data = dat, 
    fun = "event",
    ylim = c(0, 0.05),
    conf.int = TRUE,
    risk.table = TRUE,
    ggtheme = theme_minimal(),
    title = "",
    xlab = "Follow-up time (years)",
    ylab = "Cumulative incidence",
    palette = "black",
    font.main = c(14, "plain", "black"),
    font.x = c(14, "plain", "black"),
    font.y = c(14, "plain", "black"),
    font.tickslab = c(14, "plain", "black"),
    risk.table.fontsize = 4,
    break.x.by = 2)

  ## Survival (KM) plot
  ggsurvplot(
    obs, 
    data = dat, 
    ylim = c(0.95, 1),
    conf.int = TRUE,
    risk.table = TRUE,
    ggtheme = theme_minimal(),
    title = "",
    xlab = "Follow-up time (years)",
    ylab = "Survival probability",
    palette = "black",
    font.main = c(14, "plain", "black"),
    font.x = c(14, "plain", "black"),
    font.y = c(14, "plain", "black"),
    font.tickslab = c(14, "plain", "black"),
    risk.table.fontsize = 4,
    break.x.by = 2)
  
    rm(list=setdiff(ls(), c("dat_nodcis", "dat_dcishg", "dat_dcisall", "dat")))
```

## Distribution of covariates {.tabset}
### Personal, hormonal, lifestyle, and disease history
```{r 1_clinfactors}
#################### Personal Information ####################
labels <- list(
  variables=list(Age = "Age",
                 ancestry = "Race/ethnicity",
                 Height_cm = "Height (cm)",
                 Weight_kg = "Weight (kg)",
                 BMI = "Body mass index",
                 bmi_group = "Body mass index (categorized)"),
  groups=list("","Group"))

strata <- c(list(Total=dat), split(dat, dat$BC_Status))

table1(strata, labels, groupspan=c(1,2),
       render.continuous=c(.="Mean (SD)", .="Median [Q1, Q3]"),
       data=dat, topclass="Rtable1-zebra")

####################### Hormonal/Lifestyle factors ####################
labels <- list(
  variables=list(Age_Menarche = "Age at menarche",
                 Parity = "Parity",
                 age1birth = "Age at first live birth",
                 Age_Menopause = "Age at menopause",
                 MHT_use = "Hormone replacement therapy use",
                 OC_ever_used = "Oral contraceptive use",
                 gram_alcohol_per_day = "Alcohol intake per day (gram)", #overall
                 gram_alcohol_per_day2 = "Alcohol intake per day (gram) - excluding non-drinkers"),
  groups=list("","Group"))

strata <- c(list(Total=dat), split(dat, dat$BC_Status))

table1(strata, labels, groupspan=c(1,2),
       render.continuous=c(.="Mean (SD)", .="Median [Q1, Q3]"),
       data=dat, topclass="Rtable1-zebra")

###################### Personal (breast) disease history ####################
labels <- list(
  variables=list(N_Biop2 = "Number of previous breast biopsies",
                 leeft = "Age at breast cancer diagnosis",
                 oestrogeenstatus = "Estrogen receptor status",
                 progresteronstatus = "Progesterone receptor status",
                 herceptinstatus = "HER2 status",
                 omtopog = "Tumor topography",
                 omlater = "Laterality",
                 ommorf = "Morphology",
                 omgedrag = "Behavior",
                 omdiffgr = "Grade",
                 ombasisd = "Basis for diagnosis",
                 ompalpabel = "Palable",
                 ombirads = "BI-RADS classification",
                 ommultifocaal = "Multifocal Tumor"),
  groups=list("","Group"))

strata <- c(list(Total=dat), split(dat, dat$BC_Status))

table1(strata, labels, groupspan=c(1,2),
       render.continuous=c(.="Mean (SD)", .="Median [Q1, Q3]"),
       data=dat, topclass="Rtable1-zebra")
```

### Family history
```{r 1_famhist}
labels <- list(
  variables=list(family_history_BrCa = "Family history of breast cancer",
           first_degree_family_history_BrCa = "First degree family history of breast cancer",
           second_degree_family_history_BrCa = "Second degree family history of breast cancer",
           BrCA_1stDeg_relative1 = "First degree family member 1 - breast cancer",
           BrCA_1stDeg_relative2 = "First degree family member 2 - breast cancer",
           BrCA_1stDeg_relative3 = "First degree family member 3 - breast cancer",
           BrCA_2ndDeg_relative1 = "Second degree family member 1 - breast cancer",
           BrCA_2ndDeg_relative2 = "Second degree family member 2 - breast cancer",
           BrCA_2ndDeg_relative3 = "Second degree family member 3 - breast cancer",
           Age_Diag_BrCa_1stDeg_rel1 = "Age at diagnosis breast cancer - first degree family member 1",
           Age_Diag_BrCa_1stDeg_rel2 = "Age at diagnosis breast cancer - first degree family member 2",
           Age_Diag_BrCa_1stDeg_rel3 = "Age at diagnosis breast cancer - first degree family member 3",
           Age_Current_BrCa_1stDeg_rel1 = "Current age breast cancer - first degree family member 1",
           Age_Current_BrCa_1stDeg_rel2 = "Current age breast cancer - first degree family member 2",
           Age_Current_BrCa_1stDeg_rel3 = "Current age breast cancer - first degree family member 3",
           Age_Diag_BrCa_2ndDeg_rel1 = "Age at diagnosis breast cancer - second degree family member 1",
           Age_Diag_BrCa_2ndDeg_rel2 = "Age at diagnosis breast cancer - second degree family member 2",
           Age_Diag_BrCa_2ndDeg_rel3 = "Age at diagnosis breast cancer - second degree family member 3",
           Age_Current_BrCa_2ndDeg_rel1 = "Current age breast cancer - second degree family member 1",
           Age_Current_BrCa_2ndDeg_rel2 = "Current age breast cancer - second degree family member 2",
           Age_Current_BrCa_2ndDeg_rel3 = "Current age breast cancer - second degree family member 3",
           
           family_history_OvCa = "Family history of ovarian cancer",
           first_degree_family_history_OvCa = "First degree family history of ovarian cancer",
           second_degree_family_history_OvCa = "Second degree family history of ovarian cancer",
           OvCa_1stDeg_relative1 = "First degree family member 1 - ovarian cancer",
           OvCa_1stDeg_relative2 = "First degree family member 2 - ovarian cancer",
           OvCa_1stDeg_relative3 = "First degree family member 3 - ovarian cancer",
           OvCa_2ndDeg_relative1 = "Second degree family member 1 - ovarian cancer",
           OvCa_2ndDeg_relative2 = "Second degree family member 2 - ovarian cancer",
           OvCa_2ndDeg_relative3 = "Second degree family member 3 - ovarian cancer",
           Age_Diag_OvCa_1stDeg_rel1 = "Age at diagnosis ovarian cancer - first degree family member 1",
           Age_Diag_OvCa_1stDeg_rel2 = "Age at diagnosis ovarian cancer - first degree family member 2",
           Age_Diag_OvCa_1stDeg_rel3 = "Age at diagnosis ovarian cancer - first degree family member 3",
           Age_Current_OvCa_1stDeg_rel1 = "Current age ovarian cancer - first degree family member 1",
           Age_Current_OvCa_1stDeg_rel2 = "Current age ovarian cancer - first degree family member 2",
           Age_Current_OvCa_1stDeg_rel3 = "Current age ovarian cancer - first degree family member 3",
           Age_Diag_OvCa_2ndDeg_rel1 = "Age at diagnosis ovarian cancer - second degree family member 1",
           Age_Diag_OvCa_2ndDeg_rel2 = "Age at diagnosis ovarian cancer - second degree family member 2",
           Age_Diag_OvCa_2ndDeg_rel3 = "Age at diagnosis ovarian cancer - second degree family member 3",
           Age_Current_OvCa_2ndDeg_rel1 = "Current age ovarian cancer - second degree family member 1",
           Age_Current_OvCa_2ndDeg_rel2 = "Current age ovarian cancer - second degree family member 2",
           Age_Current_OvCa_2ndDeg_rel3 = "Current age ovarian cancer - second degree family member 3",
           
           family_history_PanCa = "Family history of pancreatic cancer",
           #no info (not colected in later questionnaire versions): first_degree family_history_PanCa, second_degree_family_history_PanCa
           PanCa_1stDeg_relative1 = "First degree family member 1 - pancreatic cancer",
           PanCa_1stDeg_relative2 = "First degree family member 2 - pancreatic cancer",
           PanCa_1stDeg_relative3 = "First degree family member 3 - pancreatic cancer",
           PanCa_2ndDeg_relative1 = "Second degree family member 1 - pancreatic cancer",
           PanCa_2ndDeg_relative2 = "Second degree family member 2 - pancreatic cancer",
           PanCa_2ndDeg_relative3 = "Second degree family member 3 - pancreatic cancer",
           Age_Diag_PanCa_1stDeg_rel1 = "Age at diagnosis pancreatic cancer - first degree family member 1",
           Age_Diag_PanCa_1stDeg_rel2 = "Age at diagnosis pancreatic cancer - first degree family member 2",
           Age_Diag_PanCa_1stDeg_rel3 = "Age at diagnosis pancreatic cancer - first degree family member 3",
           Age_Current_PanCa_1stDeg_rel1 = "Current age pancreatic cancer - first degree family member 1",
           Age_Current_PanCa_1stDeg_rel2 = "Current age pancreatic cancer - first degree family member 2",
           Age_Current_PanCa_1stDeg_rel3 = "Current age pancreatic cancer - first degree family member 3",
           Age_Diag_PanCa_2ndDeg_rel1 = "Age at diagnosis pancreatic cancer - second degree family member 1",
           Age_Diag_PanCa_2ndDeg_rel2 = "Age at diagnosis pancreatic cancer - second degree family member 2",
           Age_Diag_PanCa_2ndDeg_rel3 = "Age at diagnosis pancreatic cancer - second degree family member 3",
           Age_Current_PanCa_2ndDeg_rel1 = "Current age pancreatic cancer - second degree family member 1",
           Age_Current_PanCa_2ndDeg_rel2 = "Current age pancreatic cancer - second degree family member 2",
           Age_Current_PanCa_2ndDeg_rel3 = "Current age pancreatic cancer - second degree family member 3",
           
           family_history_ProsCa = "Family history of prostate cancer",
           #no info (not colected in later questionnaire versions): first_degree_family_history_ProsCa, second_degree_family_history_ProsCa
           ProsCa_1stDeg_relative1 = "First degree family member 1 - prostate cancer",
           ProsCa_1stDeg_relative2 = "First degree family member 2 - prostate cancer",
           ProsCa_1stDeg_relative3 = "First degree family member 3 - prostate cancer",
           ProsCa_2ndDeg_relative1 = "Second degree family member 1 - prostate cancer",
           ProsCa_2ndDeg_relative2 = "Second degree family member 2 - prostate cancer",
           ProsCa_2ndDeg_relative3 = "Second degree family member 3 - prostate cancer",
           Age_Diag_ProsCa_1stDeg_rel1 = "Age at diagnosis prostate cancer - first degree family member 1",
           Age_Diag_ProsCa_1stDeg_rel2 = "Age at diagnosis prostate cancer - first degree family member 2",
           Age_Diag_ProsCa_1stDeg_rel3 = "Age at diagnosis prostate cancer - first degree family member 3",
           Age_Current_ProsCa_1stDeg_rel1 = "Current age prostate cancer - first degree family member 1",
           Age_Current_ProsCa_1stDeg_rel2 = "Current age prostate cancer - first degree family member 2",
           Age_Current_ProsCa_1stDeg_rel3 = "Current age prostate cancer - first degree family member 3",
           Age_Diag_ProsCa_2ndDeg_rel1 = "Age at diagnosis prostate cancer - second degree family member 1",
           Age_Diag_ProsCa_2ndDeg_rel2 = "Age at diagnosis prostate cancer - second degree family member 2",
           Age_Diag_ProsCa_2ndDeg_rel3 = "Age at diagnosis prostate cancer - second degree family member 3",
           Age_Current_ProsCa_2ndDeg_rel1 = "Current age prostate cancer - second degree family member 1",
           Age_Current_ProsCa_2ndDeg_rel2 = "Current age prostate cancer - second degree family member 2",
           Age_Current_ProsCa_2ndDeg_rel3 = "Current age prostate cancer - second degree family member 3"),
  groups=list("","Group"))

strata <- c(list(Total=dat), split(dat, dat$BC_Status))

table1(strata, labels, groupspan=c(1,2),
       render.continuous=c(.="Mean (SD)", .="Median [Q1, Q3]"),
       data=dat, topclass="Rtable1-zebra")
```

### Breast density
```{r 1_density}
labels <- list(
  variables=list(VBD = "Volpara volumetric density %",
                 VDG4thed = "Volpara Density Grade - 4th edition",
                 VDG5thEd = "Volpara Density Grade - 5th edition"),
  groups=list("","Group"))

strata <- c(list(Total=dat), split(dat, dat$BC_Status))

table1(strata, labels, groupspan=c(1,2),
       render.continuous=c(.="Mean (SD)", .="Median [Q1, Q3]"),
       data=dat, topclass="Rtable1-zebra")
```

### Genetic data: mutations and PRS
```{r 1_genetic}
labels <- list(
  variables=list(BRCA_mutation_participant = "BRCA mutation - participant",
           BRCA_mutation_father = "BRCA mutation - father",
           BRCA_mutation_mother = "BRCA mutation - mother",
           BRCA_mutation_son = "BRCA mutation - son",
           BRCA_mutation_daughter = "BRCA mutation - daughter",
           BRCA_mutation_brother = "BRCA mutation - brother",
           BRCA_mutation_sister = "BRCA mutation - sister",
           BRCA_mutation_halfbrother = "BRCA mutation - halfbrother",
           BRCA_mutation_halfsister = "BRCA mutation - halfsister",
           BRCA_mutation_grandson = "BRCA mutation - grandson",
           BRCA_mutation_granddaughter = "BRCA mutation - granddaughter",
           #no info (not collected in any questionnaire version): third degree family mutation (e.g., maternal/paternal uncle's daughters)
           
           prs313 = "Polygenic risk score - 313 SNPs"),
  groups=list("","Group"))

strata <- c(list(Total=dat), split(dat, dat$BC_Status))

table1(strata, labels, groupspan=c(1,2),
       render.continuous=c(.="Mean (SD)", .="Median [Q1, Q3]"),
       data=dat, topclass="Rtable1-zebra")
```

### Visualize PRS distribution
```{r 1_prs_raw}
  ggplot(data=dat, aes(x=prs313, group=BC_Status, fill=BC_Status)) +
    geom_density(adjust=1, alpha=.4) +
    scale_fill_manual(values = c("#f3c483", "#5445b1")) +
    labs(title="PRS score by breast cancer status in complete cases - all 313 SNPs",
         title.position="center",
         x="PRS score - all 313 SNPs",
         y="Density") +
    theme_classic()
```

## BOADICEA scores distribution
```{r 1_boadicea}
labels <- list(
  variables=list(BOA_5yrsrisk_Model1 = "Predicted 5 year risk - Model 1",
                 BOA_5yrsrisk_Model2 = "Predicted 5 year risk - Model 2",
                 BOA_5yrsrisk_Model3a = "Predicted 5 year risk - Model 3a",
                 BOA_5yrsrisk_Model3b = "Predicted 5 year risk - Model 3b",
                 BOA_5yrsrisk_Model4 = "Predicted 5 year risk - Model 4",
                 BOA_5yrsrisk_Model5a = "Predicted 5 year risk - Model 5a",
                 BOA_5yrsrisk_Model5b = "Predicted 5 year risk - Model 5b"),
  groups=list("","Group"))

strata <- c(list(Total=dat), split(dat, dat$BC_Status))

table1(strata, labels, groupspan=c(1,2),
       render.continuous=c(.="Mean (SD)", .="Median [Q1, Q3]"),
       data=dat, topclass="Rtable1-zebra")
```

# 2-3. Data: Sensitivity analysis included DCIS
```{r 2_outcome, eval=FALSE}
# Repeat but change data
dat <- dat_dcishg
dat <- dat_dcisall
```
